Naturalness and Long-Lived Beyond the Standard Model Particles

### 2017-08-04 tex/draft.tex
Integrate Stan's comments

    Experimentally, the $\Wpm$, $\Z$, quarks, and leptons are massive. Explicit mass terms for the fermions or gauge bosons would violate the chiral and gauge symmetries of the SM Lagrangian, however, so $\SUtwoLUoneY$ must be broken\fxnote{why can't you explicitly break gauge symmetry?}. Specifically, the nonzero vacuum expectation value (VEV)\fxnote{maybe just use $v$} of the Higgs field spontaneously breaks $\SUtwoLUoneY$ to $\UoneQ$. In the process, three of the four degrees of freedom associated with the complex Higgs doublet become the longitudinal polarizations of $\Wpm$ and $\Z$, giving them mass. The remaining degree of freedom is the physical Higgs boson. Expanding the Higgs field about its VEV produces Yukawa interactions such as those in Eq.~\ref{yukawa_terms}\fxnote{This still doesn't sound right to me. The Yukawa interactions aren't produced by anything other than somebody ad-hoc adding the terms to the Lagrangian. Masses for the gauge bosons are an inherent feature of the Higgs mechanism, masses for the fermions are an ad-hoc extension, says Andrew}. These terms are interpreted as giving mass to the quarks and charged leptons, where the mass of each particle is given by the product of the Higgs VEV, $v$, and the coupling constant, $y$~\cite{sm_intro}.

@@

\section{The CMS Detector}\fxnote{Add pictures of everything}
    \subsection{Overview}
        \fxnote{define eta, coordinate system in general}
        \fxnote{Mention that ATLAS is similar, says Andrew}  
        The Compact Muon Solenoid (CMS) detector is designed to reconstruct a variety of particles from collisions at the Large Hadron Collider (LHC)\fxnote{Make sure you are able to simply explain how the LHC works during your talk: how does it accelerate protons, how does it curve their trajectories in a circle, etc, says andrew}, a circular proton-proton collider that ran with center-of-mass energy \SI{7}{\tera\electronvolt} in 2010-11, \SI{8}{\tera\electronvolt} in 2012, and \SI{13}{\tera\electronvolt} from 2015 through to the present\fxnote{cite something}. To meet this goal, CMS combines information from several subdetectors nested radially in and around a \SI{3.8}{\tesla} superconducting solenoid. All together, CMS is \SI{21.6}{\m} long, \SI{14.6}{\m} across, and weighs approximately \SI{12500}{t}~\cite{cms_experiment}. The subdetectors that make up CMS are shown in Fig.~\ref{cms} and described below.

@@

    \subsection{Electromagnetic Calorimeter}
        \fxnote{I define EE \& EB but never use those definitions}
        After traversing the inner tracker, particles next encounter the electromagnetic calorimeter (ECAL). As a homogeneous scintillation calorimeter, ECAL uses \num{61200} lead tungstate crystals in the barrel and \num{7324} in each endcap to reconstruct the energy deposited during electromagnetic showers. Lead tungstate crystals allow for a fast (\SI{80}{\percent} of light within \SI{25}{\nano\s}), compact (radiation length = \SI{0.89}{cm}), fine-grained (Moli\`ere radius = \SI{2.2}{cm}), and radiation hard (up to 10 Mrad\fxnote{learn context for this number}) calorimeter. The main drawback is the relatively low light yield (\SI{30}{photon\per\mega\electronvolt}), which necessitates photodetectors with intrinsic gain that work in magnetic fields\cite{cms_experiment, cms_tdr}.

@@

    \subsection{Muon System}
        \fxnote{Specifically mention CMS design specs to meet physics goals?}
        The muon system (MS), which is the outermost subdetector, detects muons with \num{3} varieties of gaseous ionization detectors: drift tubes (DTs) in the barrel, cathode strip chambers (CSCs) in the endcaps, and resistive plate chambers (RPCs) in both the barrel and endcaps. The choice of detector varies between barrel and endcap due to the difference in neutron-induced background\fxnote{what?} , muon rates, and magnetic field strength\fxnote{why does B matter?}. The RPCs are useful in the endcaps and barrel because their fast response and good time resolution allow unambiguous bunch crossing identification. Combining MS and tracker measurements improves the muon momentum resolution at high momenta where the tracker resolution declines \fxnote{quantify or show std plot?}\cite{cms_experiment, cms_tdr}.

@@

    \subsection{Trigger}
        \fxnote{Is entire detector read out at L1?}
        The trigger reduces the data writing rate from the \SI{40}{\mega\hertz} collision rate to less than \SI{1}{\kilo\hertz} so that events can be written to tape. The rate reduction happens in two stages: Level-1 (L1) and High-Level Trigger (HLT). L1 analyzes input from ECAL, HCAL, and MS with custom electronics to reduce the rate to $\approx$100 kHz in 3.8 us. With input from all subdetectors, the HLT then uses a dedicated processor farm to further reduce the rate to the desired $<$ 1 kHz\cite{cms_experiment, cms_trigger_upgrade}.

>>>>

    Experimentally, the $\Wpm$, $\Z$, quarks, and leptons are massive\fxnote{know about the weinberg angle, first measured precisely in neutrino electron elastic scattering}. Explicit mass terms for the fermions or gauge bosons would violate the chiral and gauge symmetries of the SM Lagrangian, however, so $\SUtwoLUoneY$ must be broken\fxnote{why can't you explicitly break gauge symmetry?}. Specifically, the nonzero vacuum expectation value (VEV)\fxnote{maybe just use $v$} of the Higgs field spontaneously breaks $\SUtwoLUoneY$ to $\UoneQ$. In the process, three of the four degrees of freedom associated with the complex Higgs doublet become the longitudinal polarizations of $\Wpm$ and $\Z$, giving them mass. The remaining degree of freedom is the physical Higgs boson. Expanding the Higgs field about its VEV produces Yukawa interactions such as those in Eq.~\ref{yukawa_terms}\fxnote{This still doesn't sound right to me. The Yukawa interactions aren't produced by anything other than somebody ad-hoc adding the terms to the Lagrangian. Masses for the gauge bosons are an inherent feature of the Higgs mechanism, masses for the fermions are an ad-hoc extension, says Andrew}. These terms are interpreted as giving mass to the quarks and charged leptons, where the mass of each particle is given by the product of the Higgs VEV, $v$, and the coupling constant, $y$~\cite{sm_intro}.

@@

\section{The CMS Detector}\fxnote{Add pictures of everything}
    \subsection{Overview}
        \fxnote{define eta, coordinate system in general}
        \fxnote{Mention that ATLAS is similar, says Andrew}  
        \fxnote{Know about the beam, says stan}  
        The Compact Muon Solenoid (CMS) detector is designed to reconstruct a variety of particles from proton-proton collisions at the Large Hadron Collider (LHC)\fxnote{Make sure you are able to simply explain how the LHC works during your talk: how does it accelerate protons, how does it curve their trajectories in a circle, etc, says andrew}, a circular collider that ran with center-of-mass energy \SI{7}{\tera\electronvolt} in 2010-11, \SI{8}{\tera\electronvolt} in 2012, and \SI{13}{\tera\electronvolt} from 2015 through to the present\fxnote{cite something}. To reconstuct final-state particles, CMS uses a particle-flow algorithm that combines information from several subdetectors nested radially in and around a \SI{3.8}{\tesla} superconducting solenoid. All together, CMS is \SI{21.6}{\m} long, \SI{14.6}{\m} across, and weighs approximately \SI{12500}{t}~\cite{cms_experiment}. The subdetectors that make up CMS are shown in Fig.~\ref{cms} and described below.

@@

    \subsection{Electromagnetic Calorimeter}
        \fxnote{I define EE \& EB but never use those definitions}
        After traversing the inner tracker, particles next encounter the electromagnetic calorimeter (ECAL). As a homogeneous scintillation calorimeter, ECAL uses \num{61200} lead tungstate crystals in the barrel and \num{7324} in each endcap to reconstruct the energy deposited during electromagnetic showers. Lead tungstate crystals allow for a fast (\SI{80}{\percent} of light within \SI{25}{\nano\s}), compact (radiation length = \SI{0.89}{cm}), fine-grained (Moli\`ere radius = \SI{2.2}{cm}), and radiation hard (up to 10 Mrad\fxnote{learn context for this number}) calorimeter. The main drawback is the relatively low light yield (\SI{30}{photon\per\mega\electronvolt}), which necessitates photodetectors with intrinsic gain that work in magnetic fields\cite{cms_experiment, cms_tdr}. Energy resolution is varies from approximately \num{1}-\SI{5}{\percent} depending on the amount of material traversed before reaching ECAL~\cite{ecal_performance}.

@@

    \subsection{Muon System}
        \fxnote{Specifically mention CMS design specs to meet physics goals?}
        The muon system (MS), which is the outermost subdetector, detects muons with \num{3} varieties of gaseous ionization detectors: drift tubes (DTs) in the barrel, cathode strip chambers (CSCs) in the endcaps, and resistive plate chambers (RPCs) in both the barrel and endcaps. The choice of detector varies between barrel and endcap due to the difference in neutron-induced background\fxnote{what?} , muon rates, and magnetic field strength\fxnote{why does B matter?}. The RPCs are useful in the endcaps and barrel because their fast response and good time resolution allow unambiguous bunch crossing identification. Combining MS and tracker measurements improves the muon momentum resolution at high momenta where the tracker resolution declines (see Fig.~\ref{muon_pt_resolution}\cite{cms_experiment, cms_tdr}.

        \noindent \begin{figure}[htbp] \begin{center}
        \includegraphics[width=0.5\textwidth]{images/muon_pt_resolution}
            \caption{Transverse momentum resolution in the tracker and muon system for small (left) and large (right) pseudorapidity~\cite{cms_image}.}
        \label{muon_pt_resolution}
        \end{center} \end{figure}

@@

    \subsection{Trigger and Particle Reconstruction}
        \fxnote{Is entire detector read out at L1?}
        The trigger reduces the data writing rate from the \SI{40}{\mega\hertz} collision rate to less than \SI{1}{\kilo\hertz} so that events can be written to tape. The rate reduction happens in two stages: Level-1 (L1) and High-Level Trigger (HLT). L1 analyzes input from ECAL, HCAL, and MS with custom electronics to reduce the rate to $\approx$100 kHz in 3.8 us. With input from all subdetectors, the HLT then uses a dedicated processor farm to further reduce the rate to the desired $<$ 1 kHz\cite{cms_experiment, cms_trigger_upgrade}.
